Stateless Model Checking with Data-Race Preemption Points

### 2015-11-13 pldi-16/di.tex
implement joshua's revisions pt1

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees.pdf}
	%\includegraphics[width=0.15\textwidth]{tree0.pdf}
	%\includegraphics[width=0.225\textwidth]{tree1.pdf}
	%\includegraphics[width=0.25\textwidth]{tree2.pdf}
	%\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only voluntary thread switches, such as {\tt yield()} or {\tt cond\_wait()}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	\label{fig:id}
\end{figure}
\section{Design}
\label{sec:design}

Motivated by state-of-the-art tools' inflexibility to change their preemption points,
Iterative Deepening searches among different combinations of such points during a test,
deciding on-the-fly whether to pursue each resulting state space or abandon it in favour of smaller ones.
Different state spaces are generated based on {\em subsets} of the preemption points which prior work would use; for example, ``preempt on all calls to {\tt mutex\_lock} but not on {\tt mutex\_unlock}''.
We show a visualization in Figure~\ref{fig:id}.

@@

For the rest of the paper, we abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC),% (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

@@

Please note the difference between data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is prone to false-positives,
we classify unprotected access pairs separately from concrete observable failures, % "concrete"?
calling such pairs {\em data race candidates}.
Should a future interleaving, preempting during those accesses,
lead to a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
%TODO: Don't do this, below. Scan paper for uses of 'false positive' where you mean to say 'benign' or 'both'.
%For brevity, we count benign data races as a subset of false positives.

@@

\newcommand\PendingJobs{\ensuremath{\mathcal{P}}}
\newcommand\SuspendedJobs{\ensuremath{\mathcal{S}}}
\newcommand\GetETA[1]{ETA(#1)}
\newcommand\GetPPSet[1]{PPSet(#1)}
With a limited CPU budget, we must avoid running tests that are likely to time out.
Hence, we separate the available PP sets into a set of {\em suspended} jobs (partially-explored state spaces with high ETAs),
and a set of {\em pending} jobs (untested ones with unknown ETAs).
When a MC instance reports an ETA too high,
we compare other pending jobs to find another one more likely to complete in time.
%
Our method for doing so, listed in Algorithm~\ref{alg:shouldworkblock}, is the heart of Iterative Deepening\footnote{
Though its worst-case performance is $O(mn)$ in the
%number of pending and suspended jobs,
sizes of $\mathcal{P}$ and $\mathcal{S}$,
in practice the non-constant portion beyond line 4 runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
Its main feature is understanding that if \GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$},
and $j_1$ is suspended,
then $j_2$'s state space is guaranteed to be strictly larger, so $j_2$ will take at least as long.
Hence we should avoid testing $j_2$ unless $j_1$'s ETA improves over time.
%reveals that it might finish in time after all.
Similarly, when some job finds a bug, we cancel all pending superset jobs, as they would find the same bug.

@@

\subsection{Classifying data races as buggy or benign}

@@

With Iterative Deepening, this is a simple matter of creating a new state space with a PP enabled on the racing instructions by each thread, as shown in Algorithm~\ref{alg:handledatarace}.
We take care to avoid duplicating any superset state spaces with PPs on multiple data races.
The new space may expose a failure, in which case we report a data-race bug,
or complete successfully, which indicates a benign or false-positive data race.
It may also uncover a new data-race candidate in some alternate interleaving,
in which case we iteratively advance to a superset state space containing PPs for both racing access pairs.
Because Iterative Deepening is constrained by a CPU budget,
we may time out before completing a data race's associated state space,
in which case it remains a potential false positive that the user must handle.

@@

We chose \landslide~\cite{landslide} as our stateless model checker due to its ability to trace program execution at the granularity of individual instructions and memory accesses, which dynamic data-race detection requires.
\landslide~implements DPOR,
%\cite{dpor},
state space estimation, and a hybrid lockset/happens-before data-race analysis.
% TODO: \cite{fairstatelessmc} instead or in addition?
It avoids state space cycles (e.g. ad-hoc synchronization with {\tt yield()} or even {\tt xchg} loops) with a heuristic similar to Fair-Bounded Search \cite{bpor}. % TODO: devote a paragraph to this, after talking about drs in locks
% this line can be cut if space is needed
It can test both user-level and kernel-level code, although is limited to timer-driven nondeterminism.
Its bug-detection metrics include assertion failure, deadlock, segfault, heap checking (like Valgrind~\cite{valgrind}), and a heuristic infinite loop/livelock check.

{\bf Restricting PPs with stack trace predicates.}
Most MCs preempt indiscriminately on any sync API call, regardless of the call-site.
However, when testing a particular module in a large codebase,
the user is likely uninterested in PPs arising from other modules.
\landslide~provides the {\tt within\_function} configuration command for a user to identify which call-sites matter most.
Before inserting a PP, \landslide~requires at least one argument to {\tt within\_function} to appear in the current thread's stack trace.
%The {\tt without\_function} directive works similarly, but as a blacklist.
The {\tt without\_function} directive is the dual of {\tt within\_function}, indicating a blacklist.
Multiple invocations can be used; later ones take precedence.
%\cite{landslide} provides further detail on this feature.

{\bf Data races in lock implementations.}
Data race tools in prior work \cite{tsan,portend} recognize the implementations of sync primitives to avoid spuriously flagging memory accesses resulting from the lock implementation itself.
Assuming the locks are already correct enables productive data-race analysis on the rest of the codebase.
With testing limited to one execution, even if you wished to test for lock bugs, data-race analysis would still uselessly flag every single access pair in the lock implementation, requiring human attention to verify.
However, Iterative Deepening can automatically verify a large quantity of data-race candidates as benign.
Hence, we extended \landslide~with a custom option to change the lock-set tracking to include accesses from {\tt mutex\_lock()} and {\tt mutex\_unlock()} in the analysis. (Accesses from other sync functions, such as {\tt cond\_wait()}, would either be included already, or be protected by an internal mutex.)
% TODO: Add a code sample showing how the lockset boundaries change? If there's room?

@@

{\bf Initial PPs.}
For testing user-space code, we seed the exploration with the four subsets of ``hard-coded'' locking API PPs shown in Figure~\ref{fig:id}:
$\{yield\}$,
$\{yield,lock\}$,
$\{yield,unlock\}$,
and $\{yield,lock,unlock\}$,
By extension, these also introduce preemptions on any other sync primitives implemented with mutexes,
such as condition variables, semaphores, and r/w locks.
Preempting on voluntary switches such as {\tt yield} is always necessary to maintain the invariant that only one thread runs between consecutive PPs.
%so the {\tt yield} PP is always implicitly enabled.

For kernel-level testing, we consider interrupt-disabling to be analogous to locking,
so we also preempt just before {\tt cli} and just after {\tt sti} (to appropriate the names of the x86 instructions)\footnote{
During data-race detection, {\tt cli}/{\tt sti} are treated as a single global lock,
as {\tt cli}'d memory accesses can still race with others that have interrupts on.
Some kernels disable preemption without disabling interrupts,
which can be modelled the same way using manual annotations of that API.}.
\quicksand~is configured to begin with the subsets
$\{yield\}$,
$\{yield,lock\}$,
$\{yield,unlock\}$,
$\{yield,cli\}$,
$\{yield,sti\}$,
and $\{yield,lock,$ $unlock,cli,sti\}$.
As a heuristic, we don't bother with every intermediate subset such as $\{lock,sti\}$,
which could potentially be improved in future work (\sect{\ref{sec:future}}).

{\bf Communication protocol.}
The interface to \landslide~, which any similar MC could implement, has two parts.
First, when starting each job, \quicksand~creates a configuration file declaring which PPs to use,
% can lose this line due to space
among other options such as mutex-testing mode,
passed as an argument to \landslide.
Then, a dedicated \quicksand~thread communicates with the \landslide~process via message-passing.%on a FIFO pipe.
\landslide~messages \quicksand~after testing each new interleaving to report updated progress and ETA,
whenever a new data-race candidate is found, and whenever a bug is found.
\quicksand~in turn replies whether the test should suspend/resume due to too high ETA, or quit due to timeout.

{\bf Heuristics.}
% List of all heuristix:
% HOMESTRETCH - last 60sec of test, don't suspend
% ETA_THRESH - "to let its ETA stabilize"
% eta factor
% shold_reproduce -- small dr jobs are not allowed to add further instances of themself (why not? don't remember)
% priority change between suspected and confirmed dr
Algorithm~\ref{alg:shouldworkblock} allows heuristically scaling a job's ETA when comparing to the time budget,
to express how optimistic we are about the estimate's accuracy.
We use 2 as this scaling factor based on the results in \cite{estimation},
though we allow changing it via the command line.
We also include a heuristic to
%ignore ETAs entirely
never suspend jobs before they pass a certain threshold of interleavings tested,
for which we choose 32,
so their ETAs have some time to stabilize.

We classify data-race candidates as {\em single-order} or {\em both-order} \cite{portend}
based on whether the MC observed the access pair ordered one or both ways while testing the original state space,
We prioritize jobs with both-order data-race PPs,
because single-order candidates are much more likely to be false-positives
(though before preempting during the access itself, we cannot say for sure, hence the heuristic).
%so we prioritize jobs with both-order data-race PPs.
For single-order races, we do not add PPs for the later access at all;
should they be needed, preempting on the first access will suffice to upgrade it to both-order.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to decide at runtime which jobs to prioritize.
% TODO: Put numbers here.

>>>>

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees.pdf}
	%\includegraphics[width=0.15\textwidth]{tree0.pdf}
	%\includegraphics[width=0.225\textwidth]{tree1.pdf}
	%\includegraphics[width=0.25\textwidth]{tree2.pdf}
	%\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only voluntary thread switches, such as {\tt yield()}. %or {\tt cond\_wait()}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	\label{fig:id}
\end{figure}
\section{Design}
\label{sec:design}

Motivated by state-of-the-art tools' inflexibility to change their preemption points,
Iterative Deepening searches among different combinations of such points during a test,
deciding on-the-fly whether to pursue each resulting state space or, or to defer it in favour of smaller ones.
Different state spaces are generated based on {\em subsets} of the preemption points which prior work would use; for example, ``preempt on all calls to {\tt mutex\_lock} but not on {\tt mutex\_unlock}''.
We show a visualization in Figure~\ref{fig:id}.

@@

For the rest of the paper, we abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC), % (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

@@

We distinguish between data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is prone to false-positives,
we classify unprotected access pairs separately from concrete observable failures, % "concrete"?
calling such pairs {\em data race candidates}.
Should a future interleaving, preempting during those accesses,
% TODO: joshua wants 'failure' to be explicitly defined
lead to a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
%TODO: Don't do this, below. Scan paper for uses of 'false positive' where you mean to say 'benign' or 'both'.
%For brevity, we count benign data races as a subset of false positives.

@@

\newcommand\PendingJobs{\ensuremath{\mathcal{P}}}
\newcommand\SuspendedJobs{\ensuremath{\mathcal{S}}}
\newcommand\GetETA[1]{ETA(#1)}
\newcommand\GetPPSet[1]{PPSet(#1)}
With a limited CPU budget, we must avoid running tests that are likely to be fruitless.
Hence, we separate the available PP sets into a set of {\em suspended} jobs (partially-explored state spaces with high ETAs),
and a set of {\em pending} jobs (untested ones with unknown ETAs).
When a MC instance reports an ETA too high,
we compare other pending jobs to find another one more likely to complete in time.
%
Our method for doing so, listed in Algorithm~\ref{alg:shouldworkblock}, is the heart of Iterative Deepening\footnote{
Though its worst-case performance is $O(mn)$ in the
%number of pending and suspended jobs,
sizes of $\mathcal{P}$ and $\mathcal{S}$,
in practice the non-constant portion beyond line 4 runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
Its main feature is understanding that if \GetPPSet{$j_1$} $\subset$ \GetPPSet{$j_2$},
and $j_1$ is suspended,
then $j_2$'s state space is guaranteed to be strictly larger, so $j_2$ will take at least as long.
Hence we should avoid testing $j_2$ unless $j_1$'s ETA improves over time.
%reveals that it might finish in time after all.
Similarly, when some job finds a bug, we cancel all pending superset jobs, as they would find the same bug.

@@

\subsection{Classifying data races as buggy or benign}
\label{sec:classifying}

@@

With Iterative Deepening, this is a simple matter of creating a new state space with a PP enabled on the racing instructions by each thread, as shown in Algorithm~\ref{alg:handledatarace}.
Note that even though a data race may involve two different instructions, we add new state spaces with only one new PP at a time.
Rather than adding a single state space configured to preempt on both involved instructions,
i.e., $AB =$ \GetPPSet{$j_0$} $\cup$ $\alpha$ $\cup$ $\beta$,
we prefer to add multiple smaller jobs which have a higher chance of completing in time. i.e.,
$A =$ \GetPPSet{$j_0$} $\cup$ $\alpha$ and
$B =$ \GetPPSet{$j_0$} $\cup$ $\beta$.
If $A$ and $B$ bug-free, they will in turn add $AB$ later.
%We take care to avoid duplicating any superset state spaces with PPs on multiple data races.
The condition on line 1 ensures that we avoid duplicating any state spaces with multiple data-race PPs;
for example, $AB$ is reachable by multiple paths through its different subsets, but should be added only once.

The new spaces may expose a failure, in which case we report a data-race bug,
or complete successfully, which indicates a benign or false-positive data race.
They may also uncover a new data-race candidate entirely, %in some alternate interleaving,
in which case we iteratively advance to a superset state space containing PPs for both racing access pairs.
%Because Iterative Deepening is
Being constrained by a CPU budget,
we may time out before completing a data race's associated state space,
in which case it remains a potential false positive that the user must handle.

@@

We chose \landslide~\cite{landslide} as our stateless model checker due to its ability to trace program execution at the granularity of individual instructions and memory accesses, which dynamic data-race detection requires.
\landslide~implements DPOR \cite{dpor},
%\cite{dpor},
state space estimation \cite{estimation}, and a hybrid lockset/happens-before data-race analysis \cite{hybriddatarace}.
% TODO: \cite{fairstatelessmc} instead or in addition?
It avoids state space cycles (e.g. ad-hoc synchronization with {\tt yield()} or even {\tt xchg} loops) with a heuristic similar to Fair-Bounded Search \cite{bpor}. % TODO: devote a paragraph to this, after talking about drs in locks
% this line can be cut if space is needed
It can test both user-level and kernel-level code, although is limited to timer-driven nondeterminism.
% joshua wants "segfault" to be "memory access error (i.e., segmentation fault, or bus error)"
Its bug-detection metrics include assertion failure, deadlock, segfault, heap checking (like Valgrind~\cite{valgrind}), and a heuristic infinite loop/livelock check.

{\bf Restricting PPs with stack trace predicates.}
Most model checkers (MCs) preempt indiscriminately on any sync API call, regardless of the call-site.
However, when testing a particular module in a large codebase,
the user is likely uninterested in PPs arising from other modules.
\landslide~provides the {\tt within\_function} configuration command for a user to identify which call-sites matter most.
Before inserting a PP, \landslide~requires at least one argument to {\tt within\_function} to appear in the current thread's stack trace.
%The {\tt without\_function} directive works similarly, but as a blacklist.
The {\tt without\_function} directive is the dual of {\tt within\_function}, indicating a blacklist.
Multiple invocations can be used; later ones take precedence.
%\cite{landslide} provides further detail on this feature.

{\bf Data races in lock implementations.}
Data race tools in prior work \cite{tsan,portend} recognize the implementations of sync primitives to avoid spuriously flagging memory accesses resulting from the lock implementation itself.
The assumption that the locks are already correct enables productive data-race analysis on the rest of the codebase.
Otherwise, with testing limited to one execution,
%even if one wishes to test for lock bugs,
data-race analysis would flag every access pair in the lock implementation, requiring human attention to verify.
However, Iterative Deepening can automatically verify a large quantity of data-race candidates as benign.
Hence, we extended \landslide~with a custom option to change the lock-set tracking to include accesses from {\tt mutex\_lock()} and {\tt mutex\_unlock()} in the analysis. (Accesses from other sync functions, such as {\tt cond\_wait()}, would either be included already, or be protected by an internal mutex.)
% TODO: Add a code sample showing how the lockset boundaries change? If there's room?

@@

{\bf Initial PPs.}
For testing user-space code, we seed the exploration with the four subsets of ``hard-coded'' locking API PPs that we showed in Figure~\ref{fig:id}:
$\{yield\}$,
$\{yield,lock\}$,
$\{yield,unlock\}$,
and $\{yield,lock,unlock\}$,
By extension, these also introduce preemptions on any other sync primitives implemented with mutexes,
such as condition variables, semaphores, and r/w locks.
Preempting on voluntary switches such as {\tt yield} is always necessary to maintain the invariant that only one thread runs between consecutive PPs.
%so the {\tt yield} PP is always implicitly enabled.

For kernel-level testing, we consider interrupt-disabling to be analogous to locking,
so we also preempt just before a disable-interrupt opcode ({\tt cli}) and just after interrupts are re-enabled ({\tt sti})\footnote{
%(to appropriate the names of the x86 instructions)
During data-race detection, {\tt cli}/{\tt sti} are treated as a single global lock.
%as {\tt cli}'d memory accesses can still race with others that have interrupts on.
Some kernels disable preemption without disabling interrupts,
which can be modelled the same way using manual annotations. %of that API.
This also assumes uni-processor scheduling; for SMP kernels, replace {\tt cli}/{\tt sti} with spinlocks.}.
\quicksand~is configured to begin with the subsets
$\{yield\}$,
$\{yield,lock\}$,
$\{yield,unlock\}$,
$\{yield,cli\}$,
$\{yield,sti\}$,
and $\{yield,lock,$ $unlock,cli,sti\}$.
As a heuristic, we don't bother with every intermediate subset such as $\{lock,sti\}$,
which could potentially be improved in future work (\sect{\ref{sec:future}}).

{\bf Communication protocol.}
The interface to \landslide~, which any similar MC could implement, has two parts.
First, when starting each job, \quicksand~creates a configuration file declaring which PPs to use,
% can lose this line due to space
among other options such as mutex-testing mode,
passed as an argument to \landslide.
Then, a dedicated \quicksand~thread communicates with the \landslide~process via message-passing. %on a FIFO pipe.
\landslide~messages \quicksand~after testing each new interleaving to report updated progress and ETA,
whenever a new data-race candidate is found, and whenever a bug is found.
\quicksand~in turn replies whether the test should suspend/resume due to too high ETA, or quit due to timeout.
We implement suspending jobs simply by making \landslide~wait on a message-passing call.
Should \quicksand~later resume a suspended job, we send a message to continue,
causing the \landslide~instance to resume exploring where it left off;
otherwise, we send a message only after time runs out, causing it to exit.

{\bf Heuristics.}
% List of all heuristix:
% HOMESTRETCH - last 60sec of test, don't suspend
% ETA_THRESH - "to let its ETA stabilize"
% eta factor
% shold_reproduce -- small dr jobs are not allowed to add further instances of themself (why not? don't remember)
% priority change between suspected and confirmed dr
Algorithm~\ref{alg:shouldworkblock} allows heuristically scaling a job's ETA when comparing to the time budget,
to express how optimistic we are about the estimate's accuracy.
We use 2 as this scaling factor based on the results in \cite{estimation},
though we allow changing it via the command line.
We also include a heuristic to
%ignore ETAs entirely
never suspend jobs before they pass a certain threshold of interleavings tested,
for which we choose 32,
so that their ETAs have some time to stabilize.

We classify data-race candidates as {\em single-order} or {\em both-order} \cite{portend}
based on whether the MC observed the racing instructions ordered one or both ways in the original state space,
Jobs with both-order data-race PPs are prioritized higher,
because single-order candidates are more likely to be false-positives
(though before preempting during the access itself, we cannot say for sure, hence the heuristic).
%so we prioritize jobs with both-order data-race PPs.
For single-order races, we do not add a PP for the later access at all;
should it be needed, preempting on the first access will suffice to upgrade the race to both-order.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence, together with the logic in \sect{\ref{sec:classifying}}, each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to prioritize the jobs at runtime.
% TODO: Put numbers here.

### 2015-11-13 pldi-16/intro.tex
implement joshua's revisions pt1

We present \quicksand, a framework for {\em Iterative Deepening} of preemption points during stateless model checking.
Named after the analogous technique in chess AI \cite{iterative-deepening-chess-ai}, our approach likewise makes progressively deeper searches of the state space until a given CPU budget is exhausted.
Rather than attempting to search a single state space with every available preemption point enabled (e.g., preempting on every pthread API call),
\quicksand~tests many different state spaces corresponding to subsets of those points, managing a model checker instance to explore each one.
It estimates the size of each state space to decide when long-running instances should be suspended, and dynamically generates new state spaces based on data race analysis.
%In fact, if given enough time to fully test all discovered data-race preemption points,
%Iterative Deepening provides a full verification of all thread schedules that could arise from preempting anywhere.
In fact, Iterative Deepening is fully general:
we have proved that if it completes all state spaces resulting from data-race preemption points,
that serves as a total verification of all possible thread interleavings of the given test program.

We evaluated \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate OS classes at Carnegie Mellon, Berkeley, and the University of Chicago.
We show that \quicksand~finds more bugs than the conventional stateless model checking approach given the same CPU budget,
and furthermore that adding data-race preemption points quickly exposes bugs missed by even the ``maximal'' state space of the conventional approach.

This paper's contributions are as follows:
\begin{enumerate}
	\item Iterative Deepening, a new technique for combining data-race analysis with stateless model checking, and \quicksand, an open source implementation of the technique;
	\item A proof of convergence, which shows that should it be possible in the given CPU budget,
		fully testing every discovered data-race preemption point is equivalent to testing all possible thread schedules;
	\item A new tactic for eliminating one class of false-positive data races,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
		%, unsound in single-pass analysis but which we prove sound when used with Iterative Deepening;
	%\item Techniques for detecting and flattening cyclic state spaces resulting from ad-hoc while-loop synchronization % TODO: is there actually room for this in the paper?
	\item A large evaluation in which \quicksand~compares favorably to stand-alone data-race detection and stateless model checking approaches.
\end{enumerate}

>>>>

We present \quicksand, a framework for {\em Iterative Deepening} of preemption points during stateless model checking.
Named after the analogous technique in chess AI \cite{iterative-deepening-chess-ai}, our approach likewise makes progressively deeper searches of the state space until a given CPU budget is exhausted.
Rather than attempting to search a single state space with every available preemption point enabled (e.g., preempting on every pthread API call),
\quicksand~tests many different state spaces corresponding to subsets of those points, managing a model checker instance to explore each one.
It estimates the size of each state space to decide when long-running instances should be suspended, and dynamically generates new state spaces based on data race analysis.
%In fact, if given enough time to fully test all discovered data-race preemption points,
%Iterative Deepening provides a full verification of all thread schedules that could arise from preempting anywhere.
In fact, Iterative Deepening is fully general:
we prove that if it completes all state spaces resulting from data-race preemption points,
that serves as a total verification of all possible thread interleavings of the given test program.

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate OS classes at Carnegie Mellon, Berkeley, and the University of Chicago.
\quicksand~finds more bugs than the conventional stateless model checking approach given the same CPU budget,
% joshua wants me to say "conventional approachES" here
and furthermore, adding data-race preemption points quickly exposes bugs missed by even the ``maximal'' state space of the conventional approach.

This paper's contributions are as follows:
\begin{enumerate}
	\item Iterative Deepening, a new technique for combining data-race analysis with stateless model checking, and \quicksand, an open-source implementation of the technique;
	\item A proof of convergence, which shows that should it be possible in the given CPU budget,
		fully testing every discovered data-race preemption point is equivalent to testing all possible thread schedules;
	\item A new tactic for eliminating one class of false-positive data races,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
		%, unsound in single-pass analysis but which we prove sound when used with Iterative Deepening;
	%\item Techniques for detecting and flattening cyclic state spaces resulting from ad-hoc while-loop synchronization % TODO: is there actually room for this in the paper?
	\item A large evaluation in which \quicksand~compares favorably to stand-alone data-race detection and stateless model checking approaches.
\end{enumerate}
